
```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

# Introduction


## Motivation

With the advent of massively parallel sequencing, millions of raw sequence reads can no be easily produced for a patient's tumor. Largely automated pipelines now exist to process these raw data, detect various types of molecular alterations (or variants), filter and review to identify high-confidence calls, and annotate these variants for functional significance. However, a major bottleneck remains at the variant interpretation stage. Genome analysts, molecular pathologists, clinical geneticists, laboratory geneticists and others are faced with a deluge of variants of potential relevance [@Good2014]. These variants must be manually reviewed and intersected with a vast ecosystem of knowledgebases and biomedical literature to provide current interpretation of their relevance for clinical application.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The variant interpretation bottleneck"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1zwvAVLWpN2mKrJLgo5ZcAuSKQrHZ8zunIYHLe-w6bBo/edit#slide=id.g22509046821_0_0")
```

An image test is below

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Major point!! example image"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1YmwKdIy9BeQ3EShgZhvtb3MgR8P6iDX4DfFD65W_gdQ/edit#slide=id.gcc4fbee202_0_141")
```

## Target Audience  

The course is intended for ...

## Curriculum  

The course covers...


```{r}
devtools::session_info()
```
