
# Resources for the Interpretation of Somatic Variants 

```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

## Learning Objectives

This chapter will cover:  

- AMP/ASCO/CAP guidelines
- ClinGen/CGC/VICC SOP
- Cancer Variant Knowledgebases and their limitations

## AMP/ASCO/CAP guidelines

This section will introduce the AMP/ASCO/CAP guidelines for the Interpretation and Reporting of Sequence Variants in Cancer [@Li2017].

### AMP Tiers and Levels

This subsection will introduce the concept of AMP Tiers and Levels

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The AMP/ASCO/CAP Standards and Guidelines classify somatic variants into four Tiers and four Levels according to their clinical significance and level of evidence"}
ottrpal::include_slide("https://docs.google.com/presentation/d/1zwvAVLWpN2mKrJLgo5ZcAuSKQrHZ8zunIYHLe-w6bBo/edit#slide=id.g22509046821_0_8")
```

## ClinGen/CGC/VICC Oncogenicity SOP

This section will introduce the Standards for the classification of pathogenicity of somatic variants in cancer (oncogenicity)[@Horak2022].

## Knowledgebases

This section will introduce relevant resources for Somatic Cancer variant interpretation including [COSMIC](https://cancer.sanger.ac.uk/cosmic){target="_blank"}, [ClinVar](https://www.ncbi.nlm.nih.gov/clinvar/){target="_blank"}, [OncoKB](https://www.oncokb.org/){target="_blank"}, others. 


