
# Introduction to the CIViC knowledgebase

```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

## Learning Objectives

This chapter will cover:  

- {Introduction to the CIViC data model}
- {Demo of the CIViC web interface}

## Introduction to the CIViC data model

For background on the CIViC resource and data model please consult the CIViC core papers [@Griffith2017; @Krysiak2022; @Krysiak2022b].

### CIViC Evidence

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The core element of CIViC is the evidence item (or EID) which includes an unstructured, human-readable evidence statement and multiple structured elements briefly describing a key piece of evidence from the biomedical literature related to the clinical interpretation of cancer variants. Evidence is linked to a molecular profile (variant) for a specific gene, disease, and optional phenotype. Evidence is further categorized into 6 types and 5 levels as shown."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1zwvAVLWpN2mKrJLgo5ZcAuSKQrHZ8zunIYHLe-w6bBo/edit#slide=id.g22509046821_0_14")
```

## Tour of the CIViC web interface

Here is a brief tour of the [CIViC v2](https://civicdb.org){target="_blank"} interface:
 
```{r, echo=FALSE}
knitr::include_url("https://www.youtube.com/embed/bpVkSP7qvgc")
```

